Spin canonical rings of log stacky curves

### 2015-07-07 curves-write-up/spin_cring.tex
edited new example

\begin{example}
\label{eg:}
In this example, we shall compute the log canonical ring of $(\sx, 3 \cdot \infty, L),$ where $\sx$ has coarse space $X$, signature $(0;3;3),$ and under the identification $X \cong \BP^1$ the closed point $\infty$ in $X$ is the point in the complement of one of the affine coordinate charts. Let the single stacky point of $\sx$ be $P \neq \infty$.

Following the technique which will later be introduced in Lemma ~\ref{lem:sat-1}, we deduce the structure of $R_L$ from the structure of the canonical ring of $R_{L_X},$ the canonical ring associated to the log spin curve of the coarse space, $(X, 3 \cdot \infty, L_X).$ Note that $R_{L_X} \cong k[x_1,x_2]$ where $\deg x_1 = \deg x_2 = 1.$ To see this, observe that by Riemann--Roch $h^0(X, L_X) = 2,$ so we will need two generators in degree 2. From the Generalized Max Noether Theorem for genus 0 curves, (see Voight
and Zureick-Brown \cite[Lemma 3.1.1]{vzb:stacky},) the multiplication maps are surjective, implying that $R_{L_X}$ is generated in degree 1. There cannot be any relations because $\proj k[x_1,x_2] \cong \proj R_{L_X} \cong X$ and $X$ is 1 dimensional. Any relation would necessarily imply that $\proj R_{L_X}$ is at most $0$ dimensional, contradicting the fact that $L_X$ is very ample, so $\proj R_{L_X} \cong X$ is $1$ dimensional.

Next, we construct the canonical ring of $R_L$ from $R_{L_X}$. By Riemann Roch, we see there is some element $y_{1,3} \in (R_L)_3$ with $\ord_{P_1}(y_{1,3}) = -1.$ We claim that there exist $a_1, a_2 \in k$ and a degree 4 polynomial $f(x_1, x_2) \in k[x_1,x_2]$ such that
\begin{align*}
	R_L \cong k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))
\end{align*}
First, note that $x_1,x_2,y_{1,3}$ generate all of $R_L$ from the Generalized Max Noether Theorem for genus 0 curves, as described in Voight
and Zureick-Brown \cite[Lemma 3.1.1]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx,L) \otimes H^0(\sx,nL) \rightarrow H^0(\sx, (n+1)L)
\end{align*}
are surjective. Next, note that a relation of the form $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)$ must exist, because by Riemann--Roch, $h^0(\sx,4L ) - h^0(\sx,4L_X) = 1,$ but $x_1 y_{1,3},x_2 y_{1,3}$ define two linearly independent elements of $h^0(\sx,4L ) \setminus h^0(\sx,4L_X).$ To complete the example, it suffices to show there are no additional relations. First, note that $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)$ is necessarily irreducible because there are no relations among $x_1, x_2$ and $y_{1,3}$ in lower degrees. Hence, it follows that $k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)) \rightarrow R_L$ is a surjection. Furthermore,
$\proj k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))$ is 1 dimensional and integral, while $\proj R_L$ is also 1 dimensional. Therefore, the surjection must be an isomorphism, as claimed.

>>>>

\begin{example}
\label{eg:adding-point}
In this example, we shall compute the log canonical ring of $(\sx, 3 \infty, L),$ where $\sx$ has coarse space $X$, $\sx$ has signature $(0;3;3),$ and $\infty$ is a fixed closed point of $X,$ different from the stacky point $P_1$ with stabilizer order 3.

Following the technique which will later be introduced in Lemma ~\ref{lem:sat-1}, we deduce the structure of $R_L$ from the structure of the canonical ring of $R_{L_X} := R(X,3 \infty, L_X).$

Note that $R_{L_X} \cong k[x_1,x_2]$ where $\deg x_1 = \deg x_2 = 1.$ To see this, observe that by Riemann--Roch $h^0(X, L_X) = 2,$ so we will need two generators in degree $1$. Because $L_X$ is very ample, $R_{L_X}$ is generated in degree 1. To conclude, note that, $R_{L_X}$ does not have any relations. If there were some relation, then $\dim \proj R_{L_X} < 1$, contradictig the fact that $L_X$ is very ample, so $\proj R_{L_X} \cong X,$ which has dimension 1.

Next, we construct generators and relations for $R_L,$ using those of $R_{L_X}$. Note that we have a natural inclusion $\iota:R_{L_X} \hookrightarrow R_L$ induced by the inclusions $H^0(\sx,L_X) \hookrightarrow H^0(\sx,L).$ By Riemann Roch, we see there is some element $y_{1,3} \in (R_L)_3$ with $\ord_{P_1}(y_{1,3}) = -1,$ not in the image of the inclusion $\iota$. We claim that there exist $a_1, a_2 \in k$ and a degree 4 polynomial $f(x_1, x_2) \in k[x_1,x_2]$ such that
\begin{align*}
	R_L \cong k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))
\end{align*}
First, note that $x_1,x_2,y_{1,3}$ generate all of $R_L$ from the Generalized Max Noether Theorem from Voight
and Zureick-Brown \cite[Lemma 3.1.1]{vzb:stacky}. That is, the maps
\begin{align*}
	H^0(\sx,3L) \otimes H^0(\sx,(n-3)L) \rightarrow H^0(\sx, nL)
\end{align*}
are surjective for $n \geq 4$. A relation of the form $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)=0$ must exist because $h^0(\sx,4L ) - h^0(\sx,4L_X) = 1,$ but $x_1 y_{1,3}$ and $x_2 y_{1,3}$ define two 
linearly independent elements with nontrivial image in the 1-dimensional vector space $H^0(\sx,4L )/ H^0(\sx,4L_X).$ So, we obtain a surjection
\begin{align}
\label{align:surjection}
k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)) \rightarrow R_L.
\end{align}
To complete the example, it suffices to show there are no additional relations. First, note that $a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2)$ is irreducible because there are no relations among $x_1, x_2$ and $y_{1,3}$ in lower degrees.
Hence,  $k[x_1,x_2,y_{1,3}]/(a_1 x_1 y_{1,3} +a_2 x_2 y_{1,3} + f(x_1,x_2))$ is integral and is 2 dimensional. Thus, the map \eqref{align:surjection} defines a surjection from an integral 2 dimensional ring to a 2 dimensional ring. Therefore, it is an isomorphism.
